---
title: "Exploratory Data Analysis"
author: "Data Sci Buds (Pranav, Lily, Minjee, Tammy)"
date: '2022-04-27'
output:
  html_document:
    toc: true
    toc_float: true
    highlight: "tango"
editor_options: 
  chunk_output_type: console
---

```{r include=FALSE}
#load packages 
library(tidyverse)
library(tidymodels)
library(janitor)
library(naniar)
library(ggplot2)
```

```{r}
#read in 
superstore_dat <- read_csv(file = "data/unprocessed/Superstore_data.csv") %>% 
  clean_names()
```

# Initial Split of Data

```{r}
# initial split 
#split data
superstore_split <- 
  initial_split(superstore_dat,
                prop = .8,
                strata = profit)

superstore_train <- training(superstore_split)
superstore_test <- testing(superstore_split)
```

```{r}
#folds
superstore_resamples <- 
  superstore_train %>% 
  vfold_cv(v = 5, repeats = 3, strata = profit)
```

# Initial Overview of Data

```{r}
#9994 observations and 21 features 
dim(superstore_dat)
```

```{r}
#check for missingness 
miss_var_table <- miss_var_table(superstore_train)
miss_var_summary <- miss_var_summary(superstore_train)
gg_miss_var <- gg_miss_var(superstore_train)
```

```{r}
#no missingness in the data 
miss_var_table
miss_var_summary
gg_miss_var
```

# Essential Findings

## Response Variables

```{r}
# univariate profit graph
ggplot(superstore_dat, aes(profit)) +
  geom_density() +
  xlim(c(-200, 300)) +
  ggtitle("Profit or Loss Incurred from Sales")
```

```{r}
# univariate sales graph
ggplot(superstore_dat, aes(sales)) +
  geom_density() +
  xlim(c(0,1000)) +
  ggtitle("Sales of Products")
```

## Important Predictor Variables

```{r}
# univariate category graph
ggplot(superstore_dat, aes(category)) +
  geom_bar() +
  ggtitle("Category of Products Sold")
```

```{r}
# univariate subcategory graph
ggplot(superstore_dat, aes(sub_category)) +
  geom_bar() +
  ggtitle("Subcategory of Products Sold") +
  coord_flip()
```

```{r}
#univariate region graph
ggplot(superstore_dat, aes(region)) +
  geom_bar() +
  ggtitle("Region where Customer is From")
```

```{r}
# univariate discount graph
ggplot(superstore_dat, aes(discount)) +
  geom_density() +
  ggtitle("Discount Provided")
```

```{r}
# univariate state graph
ggplot(superstore_dat, aes(state)) +
  geom_bar() +
  ggtitle("Customer State of Residence") +
  coord_flip()
```

```{r}
# bivariate sub-category graph
superstore_dat %>%
  group_by(sub_category) %>%
  mutate(avg_prf = mean(profit)) %>%
  ggplot(aes(sub_category, avg_prf)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(
    title = "Average Profit for Different Sub-Categories",
    x = "Subcategory",
    y = "Average Profit"
  ) +
  theme(axis.text.x = element_text(angle = 90))
```
<br>
From the graph above we can see that the copier subcategory brings exceptionally more profit than the other subcategories, with an average around 800.
<br><br>

```{r}
# same plot not including copiers  
superstore_dat %>%
  group_by(sub_category) %>%
  mutate(avg_prf = mean(profit)) %>%
  ggplot(aes(sub_category, avg_prf)) +
  geom_bar(stat="identity", position=position_dodge()) +
  scale_y_continuous(limits = c(-60, 60),
                     breaks = seq(-60, 60, 20)) +
  labs(
    title = "Average Profit for Different Sub-Categories",
    subtitle = "(Not including Copiers sub-category)",
    x = "Subcategory",
    y = "Average Profit"
  ) +
  theme(axis.text.x = element_text(angle = 90))
```
<br>
Omitting the copiers for a closer look on the other subcategories, we can see that they average around 30 in mean profit. In addition, the bookcases, supplies and tables subcategories are negative in profit, with the tables subcategory having the greatest loss of around 55. 
<br><br>

```{r}
# bivariate discount graph - profit
superstore_dat %>%
  group_by(discount) %>%
  mutate(avg_prf = mean(profit)) %>%
  ggplot(aes(discount, avg_prf)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.04) +
  scale_x_continuous(limits = c(-0.05, 0.85),
                     breaks = seq(-0.1, 0.8, 0.1),
                     expand = c(0,0)) +
  labs(
    title = "Average Profit for Different Discount Values",
    x = "Discount Value",
    y = "Average Profit"
  )
```
<br>
From the relationship between discount and profit, we can see that discount values above 0.3 generally bring a loss as opposed to a profit, with losses peaking at a discount value of 0.5. The discount value that brings the most average profit is 0.1.
<br><br>

```{r}
# bivariate discount graph - sales
superstore_dat %>%
  group_by(discount) %>%
  mutate(avg_sls = mean(sales)) %>%
  ggplot(aes(discount, avg_sls)) +
  geom_bar(stat="identity", position=position_dodge(), width = 0.04) +
  scale_x_continuous(limits = c(-0.05, 0.85),
                     breaks = seq(-0.1, 0.8, 0.1),
                     expand = c(0,0)) +
  labs(
    title = "Average Sales for Different Discount Values",
    x = "Discount Value",
    y = "Average Sales"
  )
```
<br>
Contrastingly, average sales are reflected differently with discounts, with a 0.5 discount bringing the highest average sales of over 875, and the 0.1 discount coming second highest in sales. Depending on whether the supermarket chooses to maximize profit or revenue, they can choose discount values of 0.1 for the former and 0.5 for the latter. 
<br><br>

```{r}
# bivariate table - top 10 states in sales
superstore_dat %>%
  group_by(state) %>%
  mutate(average_sales = mean(sales)) %>%
  select(state, average_sales) %>%
  arrange(desc(average_sales)) %>%
  distinct() %>%
  head(10)

# bivariate table - top 10 states in profit 
superstore_dat %>%
  group_by(state) %>%
  mutate(average_profit = mean(profit)) %>%
  select(state, average_profit) %>%
  arrange(desc(average_profit)) %>%
  distinct() %>%
  head(10)
```
<br>
Shown above are two tables showing respectively the top 10 states with the highest average sales and average profit. The two tables have a high overlap of 7 out of 10 states: Wyoming, Vermont, Rhode Island, Montana, Indiana, Missouri, and Minnesota. In particular, Vermont is highest in profit and second highest in sales.
<br><br>

```{r}
# region sales contribution and segment composition
superstore_dat %>%
  group_by(region) %>%
  mutate(avg_sls = mean(sales)) %>%
  ggplot(aes(x = region, y = avg_sls, fill = segment)) +
  geom_bar(stat="identity") +
  scale_y_continuous(limits = c(0, 750000),
                     breaks = seq(0, 800000, 100000),
                     expand = c(0,0),
                     labels = comma) +
  labs(
    title = "Region Sales and Segment Composition",
    x = "Region",
    y = "Average Sales",
    fill = "Customer Segment"
  ) + 
  theme_light()

```
<br>
This stacked bar graph shows the average sales contribution and customer segment composition of each region. We can see that in general, the consumer segment brings a larger sales contribution as opposed to corporate and home office segments, and the West region is highest in average sales.
<br><br>

```{r}
# west region state sales and segment
superstore_dat %>%
  filter(region == "West") %>%
  group_by(state) %>% 
  mutate(avg_sls = mean(sales)) %>%
  ggplot(aes(x = state, y = avg_sls, fill = segment)) +
  geom_bar(stat="identity") +
  scale_y_continuous(limits = c(0, 475000),
                     breaks = seq(0, 500000, 50000),
                     expand = c(0,0),
                     labels = comma) +
  labs(
    title = "Western States Sales and Segment Composition",
    x = "State",
    y = "Average Sales",
    fill = "Customer Segment"
  ) + 
  theme_light() +
  theme(axis.text.x = element_text(angle = 90))
```
<br>
This stacked bar graph in turn shows the average sales and customer segment composition of the Western states, where California seems to bring the highest sales contribution, with Washington as second highest. 
<br><br>

# Secondary Findings

## Standard Variable Explorations

```{r}
#univariate quantity graph
ggplot(superstore_dat, aes(quantity)) +
  geom_density() +
  ggtitle("Quantity of the Product Ordered") 
```

```{r}
# univariate ship mode graph
ggplot(superstore_dat, aes(ship_mode)) +
  geom_bar() +
  ggtitle("Order Ship Mode")  
```










